A Systematic Approach to Analyzing Syndication of Viral News Stories

### 2019-12-08 paper/paper.tex
Update on Overleaf.

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{wrapfig}
\title{A Systematic Approach to Analyzing Syndication of Viral News Stories}

@@

\begin{abstract}
In this work, we present a novel, semiautonomous methodology for analyzing syndicated news articles. We first detail our manual process of news syndication analysis from which we observe a number of unsettling patterns. From there, we go on to categorize and define two different forms of syndication: top-down and bottom-up. We then transform our manual process into a semiautomated one which combines web scraping and machine learning techniques to create and then analyze a data set of viral news stories. From the analysis, we are able to make some inferences about the nature of syndicated news articles, as well as trace back a viral news story to its original publisher and article. By presenting this work, we hope to minimize the effects that blind news syndication has on the spread of misinformation. 
\end{abstract}
\keywords{Fake news \and News propagation \and News syndication}
\newpage
\tableofcontents
\newpage
\section{Introduction}
Misinformation, disinformation, and fake news have become a common occurrence in our interconnected society. The idea of intentionally spreading fake news to sway public opinion, also known as disinformation, was brought to light by the U.S. Intelligence Committee's report on Russian meddling in the 2016 U.S. presidential election \cite{stelzenmuller2017impact}. Misinformation, or the unintentional spread of inaccurate news, has also become ubiquitous thanks to social media, which allows unreliable news to be proliferated quickly and without any verification \cite{pourghomi2017stop}. The use of a combination of misinformation and disinformation by nation states, non-state actors, and other politically motivated groups has become so prevalent that it has affected over seventy countries \cite{alba_satariano_2019}. This problem has gotten so extreme that the integrity of the U.S. election system, and even the fundamental democratic ideals of the U.S., have been called into question by some \cite{mueller2019report}.

But what exactly is fake news? According to Cambridge Dictionary \cite{FAKENEWS51:online}, fake news is "false stories that appear to be news, spread on the internet or using other media, usually created to influence political views or as a joke". In addition to this definition, fake news may also include clickbait stories, propaganda, satire, biased news, or even sometimes just bad journalism \cite{Explaine83:online}.

These days, anyone can publish content on the internet, whether it is on a website, a blog, or a social media profile, and potentially reach large audiences. The fact that people nowadays rely on the internet to consume news doesn't make this problem any easier to solve.
Fake news is very profitable for the publisher \cite{Explaine83:online}, as viral stories generate large amounts of traffic, thus increasing advertisement revenue. For that reason, fake news are widely common around the internet. 
Another motive to publish fake news is to promote a political agenda. Fake political news stories are meant to persuade consumers to accept a biased of false beliefs. This is the main reason detecting fake news is such a challenge. These false stories were created to sound true and interesting.

Due to the global nature of this problem, over fifty countries have begun to implement measures to combat misinformation and disinformation \cite{poynter}. The measures implemented by these countries span from laws and legislation, to creating task forces and launching official investigations to try to weed disinformation \cite{poynter}. Besides nation states trying to tackle this behemoth of a problem, many researchers and academics have also been trying to implement various methods to try to detect and classify misinformation, disinformation, and fake news. This area of research has become commonly known as \textit{fake news detection} \cite{zhang2019overview}.

\subsection{Problem Statement}
One can try and detect fake news by checking the source of the story, read beyond the title (which is intended to attract users and thus contains more interesting and controversial topics), try to trace the article back to the original source, or even make sure the article is not a satire by any chance \cite{Explaine83:online}.
Even though it's possible, doing so is not an easy task, and that is why there has already been many different approaches proposed to detecting and classifying fake news. Many proposed solutions have mostly relied on various machine learning techniques, such as using deep neural networks, or recurrent learning models \cite{shu2017fake}. These models for the most part work by training the model on a variety of features ranging anywhere from text-based features, to network features that show how the article has been spread through space and time \cite{shu2017fake}. The problem with these models is that they rely on a lot of information to be fed into them, which take a long time to extracted. Also, many of these approaches have only been tested on a few media outlets and social media platforms and could not work on a global scale.

@@

\subsection{Objectives}
Our first objective of this work is to build a system that given a viral article, could find related articles that were either syndicated off of the original article, or articles that were syndicated by the original article. This allows us to achieve our second objective of building a propagation path which shows where the original article was published, and by who. The third objective of this work is to then analyze the propagation path of the article and see what features change over time, how they change over time, and identify any patterns.

The last two objectives of this work were not able to be accomplished in this research, but we hope they will be completed in future work. For one, we hope to eventually fully automate this process by incorporating various models for clickbait title identification, so that the clickbait model could feed our system articles to be analyzed. Our second future objective is to eventually incorporate existing fake news detection models so that once our system identifies the original article from the propagation path, it can then be classified as fake news or not. In the end, we hope to be able to provide the media consumer with a message stating where the article originated from, and if its veracity can be verified or not.

\subsection{Contributions}
We make a number of contributions throughout this work. First, we present out initial manual methodology which outlines how we first identified viral articles, how we traced them back to their origin, and the findings from out analysis. From this analysis, we present and define two different types of news syndication: top-down and bottom-up. We then present a refined, semi-autonomous methodology which is derived from our initial methodology, but with a few revisions. This methodology, given a viral news story and a couple related articles, will automatically scrape the web and identify articles within a set time window which were either syndicated from the original set of articles fed to it, or articles that the original set of articles syndicated. We use a combination and similarity score and key words that eventually feed into a dendrogram which clusters similar articles together. We then take this clustered data set and build a timeline which allows the original article to be identified. We also are able to perform a wide range of analysis on various article features and categorize how certain features change over time. This analysis then allows us to identify patterns about syndicated articles and eventually draw conclusions.

@@

The main difference between these definitions is the intent and what it is being used for. These terms may be used interchangeably in this work. This is because our work does not intend to differentiate between the intent of the news. Our work focuses on building a methodology that is able to semi-autonomously trace back an article to its origin and analyze how article features change over time. Future work involves trying to classify news as fake or not, but that involves using already developed models for classification. Not even these models attempt to differentiate between different intentions. Therefore, although misinformation, disinformation, and fake news do technically have different meanings, they can be used interchangeably in this line of work since most research on fake news detection just attempts to classify news as accurate or inaccurate, not to try to determine the intent behind the article.

\subsection{Limitations}
As stated before, this work does not attempt to make a determination of whether the content of an article is true or false. Instead, this research is primarily focused on presenting a semiautonomous methodology which can be used to identify the propagation path of a viral article, as well as conduct analysis on article features. In the future, we hope to full automate this methodology, as well as incorporate existing fake news classification models in order to actually make a determinate about content accuracy. However, for now, the scope of this work is strictly limited to semi autonomously identifying the propagation path of an article, as well as conducting time series analysis on articles features.

@@

%The thing is, that identifying news propagation path is tedious, too. To identify the propagation path of a given article, one would have to dive deep into the results of Google search, manually identify the original source of the story, and look through many news websites and see the differences between the ways each website covers the same story. That's why we decided to try and automate this process using machine learning techniques and data analysis.
\section{Background}
The field of misinformation and fake news has become widely researched by academics ever since the Russian misinformation campaign against the 2016 United States presidential election came to light\cite{stelzenmuller2017impact}. The most straightforward approach to detecting fake news is applying various machine learning techniques, and this is what many researchers have focused on thus far.

\subsection{Fake News Detection}Shu et al. goes through a comprehensive overview of all of the various ways of attacking the problem of fake news detection and classification from a data mining perspective \cite{shu2017fake}. From the perspective of detecting fake news via machine learning techniques, Shu et al. outlines that there are generally four different types of features that can be used to detect fake news: news content features, linguistic-based features, visual-based features, and social context features. These features are then applied to one of two different types of models: news content models, or social context models . Just as they sound, news content models focus on fact checking claims in the news article or looking at stylistic features that may indicate that the article is fake. Social context models, on the other hand, focus on how users interact with the article, the virality of the article, and the propagation path of the news story. These two different methods are rarely used independently, but rather compliment each other in many hybrid approaches, such as ensemble models \cite{shu2017fake}.

\subsection{Open Issues in Fake News Detection}
Shu et al. closes by giving an overview of many of the various open areas of research in this field, as well as proposing some future research ideas. Shu et al. outlines four future research directions: data-oriented, feature-oriented, model-oriented, and application-oriented. Within the area of data-oriented research, Shu et al. brings forward a couple of issues that need to be addressed. For one, there is no large, encompassing fake news data set which can be used as a benchmark to facilitate future research \cite{shu2017fake}. In fact, this was such a large problem that in 2018, Shu et al. published another work detailing their methodology for building a large scale, open source, data set consisting of fake news, which is now available on Github \cite{shu2018fakenewsnet}. Another idea that Shu et al. proposes for future research within the data-oriented direction is early fake news detection \cite{shu2017fake}. Many of the current state-of-the-art fake news detection methods rely on data that needs to be accumulated over time and information that is not readily available when news is just starting to go viral \cite{liu2018early}. Shu et al. suggests creating a system that provides fake news alerts during the dissemination process so that misinformation can be flagged before it is viewed by a large audience. Another suggestion that is made is to look at social media posts made within some time delay of the original posts as a means of verifying the veracity of the post \cite{shu2017fake}.  
%\label{sec:headings}

\section{Related Work}
There has been work done on early fake news detection by analyzing propagation paths and time series, as well as analyzing the dynamics of news syndication. Most of this work, however, has been limited in scope in terms of the domains that it covers, as well as the type of media being analyzed. For example, past research has focused on a either a few social media platforms, a specific media outlet, or a single country \cite{liu2018early},\cite{jin2014news}, \cite{wang2009propagation}. The research that has spanned across multiple different domains of news (social media, various news outlets, etc.) has been limited in scope when considering the type of media that is being analyzed \cite{zannettou2018origins}. There has been little work, if any, done specifically on the syndication of viral news and trying to automatically analyze and track back a viral news story to its original publisher.

\subsection{Domain-Specific Approaches}
Liu et al. propose a method of early fake news detection by modeling the propagation paths as a multivariate time series. They build a classifier that uses recurrent and convolutional networks to detect fake news. They were able to achieve an accuracy of 85\% on Twitter and an accuracy of 92\% on Sina Weibo when it came to correctly classifying fake news. Liu et al. claims that they were able to detect the misinformation five minutes after it began to spread \cite{liu2018early}. The downside to this approach is that it relies on user interaction with the article to generate the data needed to feed into their classifier. This means that some users will be exposed to the fake news before it can be flagged as misinformation just due to the nature of the data set that they are using. Also, this approach has only been tested on social media platforms and micro blogs, meaning that this methodology cannot be applied on a global scale that spans hundreds of different news outlets and social media sites.

@@

Wang et al. explore and analyze the patterns of news propagation and syndication in Chinese news media. They draw comparisons between the way news diffuses through time and space and how an epidemic spreads. Something interesting that they found was that 80\% of news outlets were responsible for re-printing news articles directly from the source \cite{wang2009propagation}. Some of our findings also substantiate this claim. Like much of the other research on this topic, Wang et al. focused on only Chinese media outlets and they also did not attempt to trace back an article back to its original source without knowing what that source was in the first place. Although Wang et al.'s research brings to light some interesting patterns in the ways that news article propagate and are syndicated, the work is confined mostly to Chinese news media, which is known to be heavily regulated by the Chinese Communist Party \cite{tai2014china}.

\subsection{Media-Specific Approaches}
Zannettou et al. take a different approach to analyzing propagation; instead of looking at social media posts or news stories, they focus on politically motivated memes. Unlike any of the previous research talked about, Zannettou et al. pool data from a number of various sources: Twitter, Reddit, 4chan, and Gab. Using memes from these sites, they are able to trace the propagation path of each meme and even draw conclusions about the influence that each meme outlet has. Zannettou et al. use Hawkes process to model how memes from various sources interact with each other and quantify the influence that each meme has on others \cite{zannettou2018origins}. Although this research encompasses many different media outlets and social media platforms, it fails to analyze any other form of media besides memes. Some of the processes outlined in this work could be applied to news articles, but much of it is very specific to memes and images and cannot be generalized to news articles due to the difference in data between an image and a set of text. Nevertheless, Zannettou et al. showed that memes are often syndicated by larger media outlets which causes fringe web communities to have a much broader reach than they ever could have before \cite{zannettou2018origins}.

\section{Phase I - Manual Approach}
Our research and methodology can be split up into a manual process (phase I), and a semiautonomous process (phase II). In the manual approach, we created a data set of popular news articles and then attempted to trace them back to the original publisher. Once we identified the propagation path, we attempted to analyze how various article features changed over time.

\subsection{Dataset Creation}
We started by collecting a diverse range of articles from a number of different news sources. As seen in Appendix \ref{appendix:a}, we selected five articles from each news source and recorded each articles' associated comment count. These articles were picked out between October 5 2019 and October 8 2019. We looked for articles that were published in the past two weeks. We picked media outlets that tended to be biased either towards the political left or right as defined by the chart in Appendix \ref{appendix:b}. The reason we did not pick any mainstream, unbiased news outlets (New York Times, Washington Post, Economist, etc.) is because these news sources typically do their due diligence in ensuring that the news that they are publishing is true, and the public typically trusts these sources \cite{kearney2017trusting}. It is also worth noting that we did not consider any articles that were about, or mentioned, President Donald Trump, as these could possibly skew results. Aside from that, we did not consider the actual content of the article when picking articles for the initial dataset.

@@

\subsubsection{Propagation Analysis}
The reason this article jumped out at us is because the number of comments for this article was much higher than any article in the dataset. Once we identified this specific story, we decided to try to trace it back to its original source. The process of trying to trace an article back to its origin is an extremely tedious and time consuming task. This is because there is no systematic approach already out there that can help us gather the information that we needed.  The data used to build the propagation path for the Chinese Gold Story can be found in Appendix \ref{appendix:c}.

@@

As it turned out, the original content for the news story was actually a twitter post. What was fascinating was that none of the articles actually attributed that Twitter was the first platform where the content for the story appeared; many of the news outlets did link to the Twitter post, but not once did any of the news sources explicitly say "this content was originally posted on Twitter by a user." It is also important to notice the time delays between when the story was first posted on Twitter, and when it was picked up by mainstream media. The initial twitter post was made on September 24. It took about two days for international news outlets (PowerApple, CrimeRussina, MenaFN, Novinite) to pick up this news story. It took about seven days for mainstream news outlets (RT, Daily Star, Daily Mail, Mirror) to pick up this news articles. It is important to note these time delays as it can be used to categorize bottom-up syndication, which we will define later in this work.

\subsubsection{Article Feature Analysis}
Besides just looking at how the article traveled through space and time as it was syndicated, we were interested in looking at how the articles' properties changed over time.
\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=7cm]{paper/img/chinese-gold-article-len.png}
  \captionof{figure}{Article Length Over Time}
  \label{chinesegoldarticlelen}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=7cm]{paper/img/chinese-gold-title-len.png}
  \captionof{figure}{Title Length Over Time}
  \label{chinesegoldtitlelen}
\end{minipage}
\end{figure}

From Figures \ref{chinesegoldarticlelen} and \ref{chinesegoldtitlelen}, we can see that the general trend for both the article length and title length was that as the article was syndicated, both features became longer. By actually analyzing the content of the article, it seems as though as each article was syndicated by another news outlet, they made more inferences about the article. For example, the original Twitter post just showed a video, and short description about the video. By the time Daily Mail posted their version of the article, they were making inferences about the official's total net worth, his assets, as well as general claims about anti corruption in China. None of these topics were ever discussed by the original Twitter post. The Daily Mail just assumed these as fact and wrapped everything up into one story.

\subsection{Case Study: University of Virginia}
Once we found the story about the Chinese gold story, we wanted to see if other similar articles exhibited the same properties and characteristics. We started looking for articles which origin could be traced back to either a social media post, or a local news media outlet. One of the first articles that we identified was the story titled \textit{University of Virginia cancels 21-gun salute from Veterans Day ceremony} from Fox News.

@@

\subsubsection{Propagation Analysis}
When we traced the propagation path back to the original published article, it turned out that it was actually an opinion piece from a local newspaper which is located around the University of Virginia. Once again, we had to use a manual process to trace back the article to its origin since we started with the Fox News article, which had already been syndicated multiple times.

@@

\subsubsection{Article Feature Analysis}
We wanted to see if the general trends from the Chinese gold story of article length increasing over time and title length increasing over time would hold true with this article. If they did, this would be an interesting correlation that could be drawn between syndicated articles that originate from local news outlets, or social media platforms.

@@

\subsection{Findings}
The findings from our manual analysis are

\subsubsection{Different Types of Syndication}
Top-down and bottom-up

\subsection{Limitations of a Manual Approach}

\section{Phase II - Semiautonomous Approach }
In the semiautomated approach, we wanted to see if our findings from phase I would still hold true against a larger data set. We automated the process of tracing the propagation path and conducting analysis on the resulting articles. The resulting methodology from phase II is the main contribution of this work.

\section{Future Work}

\section{Discussion and Conclusion}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage[dvipsnames]{xcolor}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{float}
\title{A Systematic Approach to Analyzing Syndication of Viral News Stories}

@@

\begin{abstract}
In this work, we present a novel, semi-autonomous methodology for analyzing syndicated news articles. We first detail our manual process of news syndication analysis from which we observe a number of unsettling patterns. From there, we go on to categorize and define two different forms of syndication: top-down and bottom-up. We then transform our manual process into a semi-automated one which combines web scraping and machine learning techniques to create and then analyze a data set of syndicated news stories. From the analysis, we are able to make some inferences about the nature of syndicated news articles, as well as trace back a viral news story to its original publisher and article. By presenting this work, we hope to minimize the effects that blind news syndication have on the spread of misinformation. 
\end{abstract}
\keywords{Fake news \and News propagation \and News syndication}
\newpage
\tableofcontents
\newpage
\section{Introduction}
Misinformation, disinformation, and fake news have become a common occurrence in our interconnected society. The idea of intentionally spreading fake news to sway public opinion, also known as disinformation, was brought to light by the U.S. Senate Select Intelligence Committee's report on Russian meddling in the 2016 U.S. presidential election \cite{stelzenmuller2017impact}. Misinformation, or the unintentional spread of inaccurate news, has also become ubiquitous thanks to social media, which allows unreliable news to be proliferated quickly and without any verification \cite{pourghomi2017stop}. The use of a combination of misinformation and disinformation by nation states, non-state actors, and other politically motivated groups, has become so prevalent that it has affected over seventy countries \cite{alba_satariano_2019}. This problem has become so extreme that the integrity of the U.S. election system, and even the fundamental democratic ideals of the U.S., have been called into question by some \cite{mueller2019report}.

But what exactly is fake news? According to the Cambridge Dictionary, fake news is "false stories that appear to be news, spread on the internet or using other media, usually created to influence political views or as a joke" \cite{FAKENEWS51:online}. In addition to this definition, fake news may also include clickbait stories, propaganda, satire, biased news, or even sometimes just bad journalism \cite{Explaine83:online}.

These days, anyone can publish content on the internet. Whether it is on a website, a blog, or a social media profile, it can potentially reach large audiences. The fact that people nowadays rely on the internet more than ever to consume news does not make this problem any easier to solve.
Fake news is very profitable for the publisher as viral stories generate large amounts of web traffic, thus increasing advertisement revenue \cite{Explaine83:online}. For that reason, fake news is widely common around the internet. 
Another motive to publish fake news is to promote a political agenda. Fake political news stories are meant to persuade consumers to accept biased false beliefs. This is the main reason detecting fake news is such a challenge. These false stories are created to sound true and interesting.

Due to the global nature of this problem, over fifty countries have begun to implement measures to combat misinformation and disinformation \cite{poynter}. The measures implemented by these countries span everything from laws and legislation, to creating task forces and launching official investigations to try to weed out disinformation \cite{poynter}. Besides nation states trying to tackle this behemoth of a problem, many researchers and academics have also been trying to implement various methods to try to detect and classify misinformation, disinformation, and fake news. This area of research has become commonly known as \textit{fake news detection} \cite{zhang2019overview}.

\subsection{Problem Statement}
One can try and detect fake news by checking the source of the story, read beyond the title (which is intended to attract users and thus contains more interesting and controversial topics), try to trace the article back to the original source, or even make sure the article is not satirical in nature \cite{Explaine83:online}.
Even though it is possible, doing so is not an easy task, and that is why there have already been many different approaches proposed to detecting and classifying fake news. Many proposed solutions have mostly relied on various machine learning techniques, such as using deep neural networks, or recurrent learning models \cite{shu2017fake}. These models for the most part work by training the model on a variety of features ranging anywhere from text-based features, to network features that show how the article has been spread through space and time \cite{shu2017fake}. The problem with these models is that they rely on a lot of information to be fed into them, which take a long time to be extracted. Also, many of these approaches have only been tested on a few media outlets and social media platforms, so they could very well not work on a global scale.

@@

\subsection{Objectives}
Our first objective of this work is to build a system that given a viral article, can find related articles that were either syndicated off of the original article, or articles that were syndicated by the original article. This allows us to achieve our second objective of building a propagation path which shows where the original article was published, and by whom. The third objective of this work is to then analyze the propagation path of the article and see what features change over time, how they change over time, and identify any patterns.

The last two objectives of this work were not accomplished in this research, but we hope they will be completed in future work. For one, we hope to eventually fully automate this process by incorporating various models for clickbait title identification, so that the clickbait model could feed our system articles to be analyzed. Our second future objective is to eventually incorporate existing fake news detection models so that once our system identifies the original article from the propagation path, it can then be classified as fake news or not. In the end, we hope to be able to provide the media consumer with a message stating where the article originated from, and if its veracity can be verified or not.

\subsection{Contributions}
We make a number of contributions throughout this work. First, we present out initial manual methodology which outlines how we first identified viral articles, how we traced them back to their origin, and the findings from our analysis. From this analysis, we present and define two different types of news syndication: top-down and bottom-up. We then present a refined, semi-autonomous methodology which is derived from our initial methodology, but with a few revisions. This methodology, given a viral news story and a couple related articles, will automatically scrape the web and identify articles within a time window which were either syndicated from the original set of articles, or articles that the original set of articles syndicated. We use a combination of similarity score and key words that eventually feed into a dendrogram which clusters similar articles together. We then take this clustered data set and build a timeline which allows the original article to be identified. We also are able to perform a wide range of analysis on various article features and categorize how certain features change over time. This analysis then allows us to identify patterns about syndicated articles and eventually draw conclusions.

@@

The main difference between these definitions is the intent and what the information is being used for. These terms may be used interchangeably in this work. This is because our work does not intend to differentiate between the intent of the news. Our work focuses on building a methodology that is able to semi-autonomously trace back an article to its origin and analyze how article features change over time. Future work involves trying to classify news as fake or not, but that involves using already developed models for classification. Not even these models attempt to differentiate between different intentions. Therefore, although misinformation, disinformation, and fake news do technically have different meanings, they can be used interchangeably in this line of work since most research on fake news detection just attempts to classify news as accurate or inaccurate, and does not try to determine the intent behind the article.

\subsection{Limitations}
As stated before, this work does not attempt to make a determination of whether the content of an article is true or false. Instead, this research is primarily focused on presenting a semi-autonomous methodology which can be used to identify the propagation path of a viral article, as well as conduct analysis on article features. In the future, we hope to fully automate this methodology, as well as incorporate existing fake news classification models in order to actually make a determination about content accuracy. However, for now, the scope of this work is strictly limited to semi-autonomously identifying the propagation path of an article, as well as conducting time series analysis on articles features.

@@

%The thing is, that identifying news propagation path is tedious, too. To identify the propagation path of a given article, one would have to dive deep into the results of Google search, manually identify the original source of the story, and look through many news websites and see the differences between the ways each website covers the same story. That's why we decided to try and automate this process using machine learning techniques and data analysis.
\section{Background}
The field of misinformation and fake news has become widely researched by academics ever since the Russian misinformation campaign against the 2016 United States presidential election came to light \cite{stelzenmuller2017impact}. The most straightforward approach to detecting fake news is applying various machine learning techniques, and this is what many researchers have focused on thus far.

\subsection{Fake News Detection}Shu et al. go through a comprehensive overview of all of the various ways of attacking the problem of fake news detection and classification from a data mining perspective \cite{shu2017fake}. From the perspective of detecting fake news via machine learning techniques, Shu et al. outline that there are generally four different types of features that can be used to detect fake news: news content features, linguistic-based features, visual-based features, and social context features. These features are then applied to one of two different types of models: news content models, or social context models . Just as they sound, news content models focus on fact checking claims in the news article or looking at stylistic features that may indicate that the article is fake. Social context models, on the other hand, focus on how users interact with the article, the virality of the article, and the propagation path of the news story. These two different methods are rarely used independently, but rather compliment each other in many hybrid approaches, such as ensemble models \cite{shu2017fake}.

\subsection{Open Issues in Fake News Detection}
Shu et al. close by giving an overview of many of the various open areas of research in this field, as well as proposing some future research ideas. They outline four future research directions: data-oriented, feature-oriented, model-oriented, and application-oriented. Within the area of data-oriented research, Shu et al. bring forward a couple of issues that need to be addressed. For one, there is no large, encompassing fake news data set which can be used as a benchmark to facilitate future research \cite{shu2017fake}. In fact, this was such a large problem that in 2018, Shu et al. published another work detailing their methodology for building a large scale, open source, data set consisting of fake news, which is now available on Github \cite{shu2018fakenewsnet}. Another idea that Shu et al. propose for future research within the data-oriented direction is early fake news detection \cite{shu2017fake}. Many of the current state-of-the-art fake news detection methods rely on data that needs to be accumulated over time and information that is not readily available when news is just starting to go viral \cite{liu2018early}. Shu et al. suggest creating a system that provides fake news alerts during the dissemination process so that misinformation can be flagged before it is viewed by a large audience. Another suggestion that is made is to look at social media posts made within some time delay of the original posts as a means of verifying the veracity of the post \cite{shu2017fake}.

\section{Related Work}
There has been work done on early fake news detection by analyzing propagation paths and time series, as well as analyzing the dynamics of news syndication. Most of this work, however, has been limited in scope in terms of the domains that it covers, as well as the type of media being analyzed. For example, past research has focused on either a few social media platforms, a specific media outlet, or a single country \cite{liu2018early},\cite{jin2014news}, \cite{wang2009propagation}. The research that has spanned across multiple different domains of news (social media, various news outlets, etc.) has been limited in scope when considering the type of media that is being analyzed \cite{zannettou2018origins}. There has been little work, if any, done specifically on the syndication of viral news and trying to automatically analyze and trace back a viral news story to its original publisher.

\subsection{Domain-Specific Approaches}
Liu et al. propose a method of early fake news detection by modeling the propagation paths as a multivariate time series. They build a classifier that uses recurrent and convolutional networks to detect fake news. They were able to achieve an accuracy of 85\% on Twitter and an accuracy of 92\% on Sina Weibo when it came to correctly classifying fake news. Liu et al. claim that they were able to detect the misinformation five minutes after it began to spread \cite{liu2018early}. The downside to this approach is that it relies on user interactions with the article to generate the data needed to feed into their classifier. This means that some users will be exposed to the fake news before it can be flagged as misinformation just due to the nature of the dataset that they are using. Also, this approach has only been tested on social media platforms and micro blogs, meaning that this methodology cannot be applied on a global scale that spans hundreds of different news outlets and social media sites.

@@

Wang et al. explore and analyze the patterns of news propagation and syndication in Chinese news media. They draw comparisons between the way news diffuses through time and space and how an epidemic spreads. Something interesting that they found was that 80\% of news outlets in China were responsible for re-printing news articles directly from the source \cite{wang2009propagation}. Some of our findings also substantiate this claim. Like much of the other research on this topic, Wang et al. focus only on Chinese media outlets and they also did not attempt to trace an article back to its original source without knowing what that source was in the first place. Although Wang et al.'s research brings to light some interesting patterns in the ways that news articles propagate and are syndicated, the work is confined mostly to Chinese news media, which is known to be heavily regulated by the Chinese Communist Party \cite{tai2014china}.

\subsection{Media-Specific Approaches}
Zannettou et al. take a different approach to analyzing propagation; instead of looking at social media posts or news stories, they focus on politically motivated memes. Unlike any of the previous research discussed, Zannettou et al. pool data from a number of various sources: Twitter, Reddit, 4chan, and Gab. Using memes from these sites, they are able to trace the propagation path of each meme and even draw conclusions about the influence that each meme outlet has. Zannettou et al. use Hawkes process to model how memes from various sources interact with each other and quantify the influence that each meme has on other memes \cite{zannettou2018origins}. Although this research encompasses many different media outlets and social media platforms, it fails to analyze any other form of media besides memes. Some of the processes outlined in this work could be applied to news articles, but much of it is very specific to memes and images and cannot be generalized to news articles due to the difference in data between an image and a set of text. Nevertheless, Zannettou et al. showed that memes are often syndicated by larger media outlets which causes fringe web communities to have much broader reach than they ever could have before \cite{zannettou2018origins}.

\section{Phase I - Manual Approach}
Our research and methodology can be split up into a manual process (phase I), and a semi-autonomous process (phase II). In the manual approach, we created a data set of popular news articles and then attempted to trace them back to the original publisher. Once we identified the propagation path, we attempted to analyze how various article features changed over time. Although inefficient, the takeaways from phase I of our research helped inform our much more efficient and effective methodology in phase II.

\subsection{Dataset Creation}
We started by collecting a diverse range of articles from a number of different news sources. As seen in Appendix \ref{appendix:a}, we selected five articles from each news source and record each articles' associated comment count. These articles were picked between October 5 2019 and October 8 2019. We looked for articles that were published in the past two weeks. We picked media outlets that tended to be biased either towards the political left or right as defined by the chart in Appendix \ref{appendix:b}. The reason we did not pick any mainstream, unbiased news outlets (New York Times, Washington Post, Economist, etc.) is because these news sources typically do their due diligence in ensuring that the news that they are publishing is true, and the public typically trusts these sources \cite{kearney2017trusting}. It is also worth noting that we did not consider any articles that were about, or mentioned, President Donald Trump, as these could possibly skew results. Aside from that, we did not consider the actual content of the article when picking articles for the initial dataset.

@@

\subsubsection{Propagation Analysis}
The reason this article jumped out at us is because the number of comments for this article was much higher than any article in the dataset. Once we identified this specific story, we decided to try to trace it back to its original source. The process of trying to trace an article back to its origin is an extremely tedious and time consuming task. This is because there is no systematic approach already out there that can help us gather the information that we need.  The data used to build the propagation path for the Chinese Gold Story can be found in Appendix \ref{appendix:c}.

@@

As it turns out, the original content for the news story was actually a twitter post. What is fascinating is that none of the articles actually state that Twitter was the first platform where the content for the story appeared; many of the news outlets did link to the Twitter post, but not once did any of the news sources explicitly say "this content was originally posted on Twitter by a user." It is also important to note the time delays between when the story was first posted on Twitter, and when it was picked up by mainstream media. The initial twitter post was made on September 24. It took about two days for international news outlets (PowerApple, CrimeRussina, MenaFN, Novinite) to pick up this news story. It took about seven days for mainstream news outlets (RT, Daily Star, Daily Mail, Mirror) to pick up this news articles. It is important to note these time delays as they can be used to categorize bottom-up syndication, which we will define later in this work.

\subsubsection{Article Feature Analysis}
Besides just looking at how the article travels through space and time as it is syndicated, we were interested in looking at how the articles' properties change over time.
\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=5cm]{paper/img/chinese-gold-article-len.png}
  \captionof{figure}{Article Length Over Time}
  \label{chinesegoldarticlelen}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=5cm]{paper/img/chinese-gold-title-len.png}
  \captionof{figure}{Title Length Over Time}
  \label{chinesegoldtitlelen}
\end{minipage}
\end{figure}

From Figures \ref{chinesegoldarticlelen} and \ref{chinesegoldtitlelen}, we can see that the general trend for both the article length and title length is that as the article is syndicated, both features become longer. By actually analyzing the content of the article, it seems as each article is syndicated by another news outlet, they make more inferences about the article. For example, the original Twitter post just shows a video, and short description about the video. By the time Daily Mail posts their version of the article, they are making inferences about the official's total net worth, his assets, as well as general claims about anti corruption in China. None of these topics are ever discussed by the original Twitter post. The Daily Mail just assumes these as fact and wrap everything up into one story.

\subsection{Case Study: University of Virginia}
Once we found the story about the Chinese gold story, we wanted to see if other similar articles would exhibit the same properties and characteristics. We started by looking for articles whose origin could be traced back to either a social media post, or a local news media outlet. One of the first articles that we identified was the story titled \textit{University of Virginia cancels 21-gun salute from Veterans Day ceremony} from Fox News.

@@

\subsubsection{Propagation Analysis}
When we traced the propagation path back to the original published article, it turned out that it was actually an opinion piece from a local newspaper which is located around the University of Virginia. Once again, we have to use a manual process to trace back the article to its origin since we started with the Fox News article, which had already been syndicated multiple times.

@@

\subsubsection{Article Feature Analysis}
We wanted to see if the general trends from the Chinese gold story of article length increasing over time and title length increasing over time would hold true with this article. If they did, this could lead to an interesting correlation that could be drawn between syndicated articles that originate from local news outlets, or social media platforms.

@@

\subsection{Findings}
Since our manual approach is extremely tedious and time consuming, the data that can be used to draw conclusions is quite limited. However, the findings from the manual approach help us better inform our phase II methodology. We are also able to make some hypothesises about patterns in syndicated news articles which are then either substantiated or disproved by the findings from phase II of our research. Also, using our observations from phase I, we are able to characterize and define two different types of news syndication.

\subsubsection{Patterns and Trends}
One of the first patterns that we observed was that syndicated articles are generally not syndicated immediately; there is typically a delay of about five to seven days from when the article is first posted and when a mainstream media outlet syndicates it. We also learned that the article length and title length of syndicated articles are always longer than that of the original article. Upon manual inspection of syndicated articles, we found that they rarely, if ever, referenced that their article was a syndication of another article. They did sometimes cite where they got their data from, but ironically, this was usually just another syndicated article. This is why tracing back a syndicated article back to its origin is such a tedious and time consuming task.
